---
title: "rapR"
subtitle: "Taking a look at the relationship between rapper net worth and money lyrics"
author: "Jake Weinreb"
output: 
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

Rapping.  It's one of the only professions I know of where you can literally make money by talking about how much money you make.  

I was thinking about this while driving one weekend, and Drake's *In My Feelings* comes on the radio.  For those of you that aren't familiar, it was a Summer 2018 hit that spawned the viral #InMyFeelingsChallenge:

![](https://www.youtube.com/watch?v=DRS_PpOrUZ4)

On the track, Yung Miami of hip-hop duo City Girls has a line that goes, "F*** that Netflix and chill what's your net-net-net worth?"

Couldn't have said it better myself, Yung Miami.

So I decided to check the receipts (aka data).  How often do rappers rap about money, what synonyms do they use to do so, and is there any relationship between how much they have and how much they talk about having?

```{r libraries}
library(spotifyr)
library(tidyverse)
library(datapasta)
library(googlesheets)
library(knitr)
```

```{r environment}
Sys.setenv(SPOTIFY_CLIENT_ID = 'my_id') #see the spotifyr package details for setup instructions
Sys.setenv(SPOTIFY_CLIENT_SECRET = 'my_secret')
access_token <- get_spotify_access_token()
```

To get the data, I used the spotifyr package, which is an R wrapper (no pun intended) for the Spotify API [^1].

[^1:] h/t Charlie Thompson, you can find more information about the package [here](https://www.rcharlie.com/spotifyr/index.html)

I grabbed the full discography, including every album, song, and lyric, for the most popular rappers today and a variety of rappers from the 80's and 90's based on public playlists.  This gave me a list of over 130 artists, from 2Pac and Lil' Kim, to 2 Chainz and Lil Uzi Vert.

```{r today}
#get top 100 most popular rappers today, filter to ones that can be considered pure rap
top_50 <- get_genre_artists("rap", limit = 50) 
next_50 <- get_genre_artists("rap", offset = 50, limit = 50)
current_rappers <- bind_rows(top_50, next_50) %>%
  unnest(genres) %>%
  filter(genres == "rap") %>% 
  select(artist_name = name, artist_id = id, artist_genre = genres, artist_followers = followers.total)
```

```{r}
#get popular artists from past
i_love_90s <- get_playlist_tracks("37i9dQZF1DX186v583rmzp") %>%
  select(track.artists) %>% 
  unlist() %>% 
  .[str_detect(names(.), pattern = "track.artists.id")] %>% 
  unname() %>% 
  unique()

i_love_80s <- get_playlist_tracks("37i9dQZF1DX2XmsXL2WBQd") %>%
  select(track.artists) %>% 
  unlist() %>% 
  .[str_detect(names(.), pattern = "track.artists.id | track.artists.name")] %>% 
  unname() %>% 
  unique()

old_rappers_info <- c(i_love_80s, i_love_90s) %>%
  unique()
```

```{r}
#loop over old_rappers to get genre and popularity info on each one
old_rappers_raw <- map(old_rappers_info, get_artist)
output <- vector("list", length = length(old_rappers_info))
for(i in seq_along(old_rappers_info)) {
  output[[i]]$artist_name <- old_rappers_raw[[i]]$name
  output[[i]]$artist_id <- old_rappers_raw[[i]]$id
  output[[i]]$artist_genre <- old_rappers_raw[[i]]$genres
  output[[i]]$artist_followers <- old_rappers_raw[[i]]$followers$total
}
old_rappers <- data.table::rbindlist(output) %>% 
  filter(artist_genre == 'rap') %>% #eliminate old artists that aren't pure "rap"
  mutate(artist_genre = as.character(artist_genre))
```

```{r}
#combine old school and new school rappers
all_artists <- bind_rows(old_rappers, current_rappers) %>%
  .$artist_name %>%
  unique()
```

```{r}
#get the full discography of all artists (this takes a while)
start_time <- Sys.time() #keep track of time
discog_raw <- map(all_artists, possibly(get_discography, otherwise = NULL))
end_time <- Sys.time()
elapsed <- end_time - start_time
discog <- bind_rows(discog_raw)
```

I also found a couple online lists of rap synonyms for the word "money"[^2].  You can check them out in the Google Sheet I included in the Github repo.

Side note: rappers love to use food terms to describe their earnings, from bacon and lettuce, to zucchini and cheese.  If I have any chef/rappers in my dataset, it may throw the analysis off!  (No, Action Bronson is not in here.)  

[^2]: The list is a bit messy, as some of the synonyms may be used in other contexts.  If you have any suggestions for additions, let me know!

```{r money_words}
#load in synonyms of money (used datapasta for this)------------------------------
money <- c("Bacon", "• Bank", "• Beans", "Benji", "• Benjamins", "• Biscuits", "• Big Ones", "Billionaire", 
           "• Bills", "• Bisquick", "• Bits", "• Boffo", "• Bones", "• Boodle", "• Bread", "• Brisket", "• Broccoli", 
           "• Bucks", "• Bullets", "• Bundle", "• Cabbage", "• Cake", "• Capital", "• Cash", "• Cheddar", "• Cheese", 
           "• Chips", "• Chits", "• Chump Change", "• Clams", "• Cnotes", "• Coin", "• Collard Greens", "• Cream", 
           "• Currency", "• Dead Presidents", "• Dibs", "• Din", "Dollar", "• Doubloons", "• Dough", "• Drafts", 
           "• Ducats", "• Duckets", "• Dub", "• Fetti", "• Fetta", "• Finances", "• Fins", "• Fish Sticks", 
           "• Fold Ups", "• Fortune", "• Franklins", "Frog", "• Frogskins", "• Funds", "• Gs", "• Gelt", "• Gold", 
           "Grand", "Green", "Greenback", "• Guala", "• Guap", "Gwop", "• Gwap", "• Guineas", "• Hundies", "Hunnit", 
           "• Jack", "• Jacksons", "• Kings ransom", "• Knots", "• Lettuce", "• Lincolns", "• Long green", "• Loonies", 
           "• Loot", "• Lucci", "• Luchini", "• Lucre", "• Mazuma", "• Means", "Millie", "Millionaire", "• Mint", 
           "• Mite", "• Moolah", "• Moss", "• Mozzarella", "• Notes", "• Paper", "• Payola", "• Pelf", "peso", 
           "piece of eight", "• Pieces of eight", "• Pin", "• Pittance", "• Plaster", "• Ps", "• Quiche", "• Quid", 
           "• Racks", "• Reagans", "• Rich", "• Rivets", "• Sawbucks", "• Scratch", "• Scrilla", "• Scrip", "• Shekels", 
           "• Shrapnel", "• Silver", "• Simoleons", "• Singles", "• Smack", "• Sodium", "• Specie", "• Spinach", 
           "• Spondulicks", "• Stacks", "• Stake", "• Stash", "• Tender", "• Treasure", "• VBucks", "• Wad", 
           "• Wampum", "Yard", "• Zucchini") %>%
  str_replace_all("[^[:alnum:]]", "")
```

```{r}
artists_clean <- discog %>%
  ungroup() %>%
  group_by(artist_name) %>%
  distinct(track_name, .keep_all = TRUE) %>% #make sure there are no duplicate track names within each artist
  filter(!map_lgl(lyrics, is.null)) %>%
  select(artist_name, artist_id, album_name, album_id, album_release_year, duration_ms, track_name, track_uri, 
         lyrics) %>%
  unnest(lyrics) %>%
  ungroup() %>%
  mutate(lyric = tolower(str_replace_all(lyric, "[:punct:]", ""))) %>% #clean capitalization, punctuation
  mutate(money_count = str_count(.$lyric, 
                           pattern = paste("\\b", tolower(money),"\\b", sep="", collapse = "|"))) %>% #add counts of money words
  filter(!str_detect(tolower(track_name), "- live")) #remove song duplicates from live albums
```

```{r}
#check which artists were cut out
artists_clean_unique <- artists_clean$artist_name %>% unique()
setdiff(all_artists, artists_clean_unique)
```

The first thing I wanted to check is how the use of money words changed over time.  Were older artists more/less likely to rap about their cash?

The chart below shows that usage has fluctated throughout time, and has been going up for several years.

```{r over_time}
use_over_time <- artists_clean %>% 
  rowwise() %>% #slower because the mutate won't use vectorization
  mutate(n_words = str_split(lyric, pattern = " ") %>% unlist() %>% length()) %>% 
  group_by(artist_name, album_name, album_release_year, track_name) %>% 
  summarize(num_words = sum(n_words, na.rm = TRUE),
            count_ratio = sum(money_count, na.rm = TRUE) / num_words) %>% 
  ungroup()
```

```{r over_time_plot}
#plot use of money words over time--maybe make this a gganimate?
use_over_time %>% 
  group_by(album_release_year) %>% 
  summarize(mean_count_ratio = mean(count_ratio),
            med_count_ratio = median(count_ratio)) %>% 
  ggplot(aes(x = album_release_year, y = mean_count_ratio)) + #could add size = popularity
  geom_point() +
  geom_smooth(method = "loess", se = FALSE) +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal() +
  labs(title = "\"I got all this money in my [lyrics] and that s*** is goin up\"--Iamsu!", 
       subtitle = "Synonyms for money as a percentage of total song lyrics",
       x = "Release Year", y = "Percent")
```

Among this group of rappers, Gucci Mane reigns supreme in his money references.  And it's not even close.  

Whether calculated by total references or percentage of lyrics, Gucci owns this lyrical theme.

```{r}
#gucci mane uses by far the most money words (percent)
use_over_time %>% 
  arrange(-count_ratio) %>% 
  filter(count_ratio > .03) %>% 
  count(artist_name) %>% 
  arrange(-n) %>% 
  head(10) %>% 
  kable()
```

```{r}
#gucci mane uses by far the most money words (count)--NEEDS UPDATING
use_over_time %>% 
  arrange(-count_ratio) %>% 
  filter(count_ratio > .03) %>% 
  count(artist_name) %>% 
  arrange(-n) %>% 
  head(10) %>% 
  kable()
```

I guess that's not really surprising, given that he has `<r artists_clean %>% filter(artist_name == "Gucci Mane") %>% .$track_name %>% unique() %>% tolower() %>% str_count("money") %>% sum()>` songs where the **title** includes a synonym for money.  In his autobiography he writes, "As far back as I can remember, I really just wanted to get me some money".  The man's not shy about his goals!

![](https://www.youtube.com/watch?v=RLFCmXGe30c)

Next we can look at which synonyms are used.  Using the full dataset, the wordcloud below gives a quick glance at the most commonly used words.  

```{r wordcloud}

```

Over time, we can see that the most popular phrasing has fluctuated.   

```{r synonyms}
#can look into which money words are used by which rappers and overall
artists_unnested <- artists_clean %>% tidytext::unnest_tokens(word, lyric)
artists_unnested %>% mutate(money_lgl = ifelse(word %in% tolower(money), 1, 0)) %>% 
  filter(money_lgl == 1) %>% 
  group_by(album_release_year) %>% 
  count(word) %>% 
  top_n(1, n) %>% 
  arrange(album_release_year, n) %>% 
  kable()
```

And here's a look at the three most popular synonyms for a few of my favorite artists:

```{r synonyms_fn}  
artist_top_money_words <- function(artist, k = 1){
  artists_unnested %>% mutate(money_lgl = ifelse(word %in% tolower(money), 1, 0)) %>% 
    filter(money_lgl == 1, artist_name == artist) %>% 
    group_by(album_release_year) %>% 
    count(word) %>% 
    top_n(k, n) %>% 
    arrange(album_release_year, n)
}

artist_synonyms <- c("Kendrick Lamar", "A$AP Rocky", "ScHoolboy Q")
map(artist_synonyms, artist_top_money_words, k = 3)
```



```{r}
#create list of rappers to gather net worth data on
#consider changing to rappers with at least 1 album in last 5 years and no albums before 2000
#maybe consider using song popularity as a proxy for sales?
net_worth_list <- artists_clean %>% 
  filter(album_release_year > "2011-01-01") %>%
  select(artist_name, album_release_year) %>% 
  group_by(artist_name) %>% 
  distinct(album_release_year, .keep_all = TRUE) %>% 
  summarize(num_albums = n()) %>% 
  filter(num_albums > 2)
```

```{r}
#full list for net worth data...more data
net_worth_list <- artists_clean %>% 
  select(artist_name) %>% 
  unique()
```

```{r}
gs_title("rapR") %>% gs_ws_new(ws_title = "Net Worth 2", input = net_worth_list) #have to log in if not already
```

```{r}
#load in net worth data from google spreadsheet
net_worth_df <- gs_title("rapR") %>% 
  gs_read(ws = "Net Worth 2") %>% 
  mutate(net_worth = str_remove_all(net_worth, ",") %>% as.numeric())
```

```{r}
#plot frequency of money words vs. net worth
use_over_time %>% 
  group_by(artist_name) %>% 
  summarize(count_ratio = mean(count_ratio)) %>% 
  full_join(net_worth_df) %>% 
  ggplot(aes(net_worth, count_ratio)) + 
  geom_point() + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  scale_x_log10()
```





